---
title: "accumulate_python"
output: html_document
---

Users, please note that this process can only be run if the shade_model_preprocessing.Rmd
process has been run through the BFW calculation step, and a HUC4 geodatabase was
downloaded as a part of that process. 

```{r}
annual_precip <- get_annual_precip(gdb_path, nhdphr_lines)

relationship <- get_relationship_table(gdb_path, nhdphr_lines)

write_csv(annual_precip, "python/annual_precip.csv")
write_csv(relationship, "python/relationships.csv")
```

If you wish to accumulate a variable other than precipitation, you can upload a 
dataset and corresponding relationship table containing the NHDPlusIDt column and
a column containing the corresponding value you wish to accumulate. In the /python
folder of this R project, upload your csv and change the names of the variables
below to match your files and the column you wish to accumulate. Please note that
the input data must have a field called NHDPlusIDt and the relationship csv must
contain the columns "FromNHDPID" and "ToNHDPID".

```{r}
metric <- "PrecipA"
input_data <- "annual_precip.csv"
input_relshpCSV <- "relationships.csv"
```

```{r}
source_python("python/accumulatePrecip.py")
accumulate_precip(metric, input_data, input_relshpCSV)
```

If a python package has not been installed in this environment, you can use the 
function below to install it. Pandas may need to be installed the first time you 
run this tool. 

```{r}
py_install("pandas")
```

